On the properties of discs around accreting brown dwarfs

### 2010-06-10 BD_disc/paper.tex
Newest Inner edge Figures
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@2290 25464f21-6099-402c-b30a-640f1d40d38b
removed abstract stuff
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@2291 25464f21-6099-402c-b30a-640f1d40d38b
Almost version
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@2294 25464f21-6099-402c-b30a-640f1d40d38b

\begin{abstract}
To Do
\begin{enumerate}
\item{Read new papers and put in any relevant stuff.}
\item{Sort structure paragraph at end of intro.}
\end{enumerate}
\end{abstract}

@@

Comparison of accretion rates across the pre-main-sequence (PMS) mass
spectrum has indicated that the accretion rate is strongly correlated
with PMS mass,
\citep{muzerolle_2003,natta_2004,natta_2006},
with an approximate form $\dot{M_{\rm acc}}\propto M_{*}^{~2}$. Since
in the canonical picture the accretion rate is driven by the disc
viscosity, and should be indepedent of the mass of the central object,
this correlation is somewhat surprising. There is some danger that the
correlation is the result of, or at least strengthened by, the
presence of observational biases \citep{clarke_2006}: At the
high-mass (CTTS) end of the mass spectrum the lowest accretion-rates
cannot be measured via continuum methods since the excess is too small
in contrast with the photopsheric emission. The emission lines will
also be weak, and indeed the H$\alpha$ equivalent width (EW) may be
less than the 10\AA\ that traditionally demarcates classical from
weak-lined T Tauri stars. Such objects should show doppler-broadened
profiles, but the line wings will be weak, and the presence of
underlying H$\alpha$ absorption may become dominant. Thus there may be
a population of low-accretion-rate CTTS which are current missing from
the surveys.

@@

Nonetheless it is worth considering the observed effect that mass
accretion at typical CTTS rates onto a BD might have. Parity between
the BD photospheric luminosity and the accretion luminosity occurs at
relatively low accretion rates \citep{clarke_2006}: Such objects may
not be detected as BDs at all in photometric surveys. Even before such
an extreme case is reached, the additional luminosity provided by the
accretion should have measurable effect on the BD colours. Furthermore
the additional flux will be reprocessed by the disc, altering its
scaleheight and possibly the shape of its inner edge.

@@

Little observational evidence can be found for approximate sizes of
accretion hot spots due to their more transient nature and often
smaller coverages, when compared to cooler or `plage' spots
\citep{herbst_2007}. \cite{bouvier_1995} modeled the
size of the cool spots on solar-type stars for a selection of
periodically variable candidates. They found projections of cooler
spots, onto the stellar disc, of a few to $~$60\%.
\cite{bouvier_1995} also found projected sizes, onto the
stellar disc, of typically a few \% to around 10\% for hot spots.
\cite{bertout_1996} used observations of YY Orionis monitoring
flux amplitude variations as a function of wavelength to derive a
probable hot spot area of around 10\%. The spot temperature was also
modeled for YY Orionis in \cite{bertout_1996}, resulting in a
best fitting areal coverage of 11\%. Therefore, to bound the probable
areal coverage range of the accretion hot spots we have adopted areal
coverages of 1 and 10\%.

@@

Accretion rates derived for pre-MS stars are of order \logmdot = $-6$
to $-11$ \citep{natta_2006}, with the largest accretion rates found in
so-called FU Orionis type objects. For the more typical accretion
rates \citep[$\dot{M}=$10$^{-11}$ to 10$^{-8.9} M_{\odot}yr^{-1}$, for
  TTS,][]{dahm_2009}, several studies have now suggested that the
accretion rate is strongly correlated with the mass of the central
star. This relationship was perhaps first suggested by
\cite{muzerolle_2003} using various accretion diagnostics. Later,
\citep{muzerolle_2005} derived a relationship of approximately
$\dot{M}\propto M_*^{~2}$. Further evidence was put forward by
\cite{natta_2004}, where accretion rates as low as 5$\times$
10$^{-12}M_{\odot} yr^{-1}$ were found for BD stars. More recently,
even lower accretion rates of $\approx$10$^{-13}M_{\odot} yr^{-1}$,
have been derived for BDs by \cite{herczeg_2009}.  Further support for
a dependence of accretion rate on stellar mass was apparent in the
significantly more homogeneous dataset of
\cite{natta_2006}. \cite{natta_2006} analysed a set of accretion rates
and masses derived for pre-MS stars in $\rho$ Ophiuchi and compared
these results to stars in Taurus. They found that the accretion rate
scales with central object mass into the BD regime, although with
significant scatter.

As the relationship $\dot{M}\propto M_*^{~2}$ predicts lower accretion
rates for BD mass objects it is essential that we model systems at
higher accretion rates, which may have been missed in current
observational studies. Therefore, we have adopted accretion rates of
\logmdot = $-6$, $-7$, $-8$, $-9$, $-10$, $-11$ \& $-12$.

@@

The initial density structure has the the disc surface density
conserved, $\Sigma ^{\beta - \alpha}$. The initial scaleheight,
$h\propto r^{\beta}$ ($h=h_0(r/R_*)^{\beta}$ where h=scaleheight and
r=radial coordinate), where r is the radial distance from the star,
and the density, $\rho \propto r^{-\alpha}$ ($\rho
=\rho_0\frac{R_*}{r}exp(-\frac{1}{2}[z/h(r)]^2)$ where h=scaleheight,
r=radial coordinate and z=vertical coordinate). The initial values of
$\alpha$ and $\beta$ were 2.1 and 1.1 respectively. The values for
$\alpha$ and $\beta$ were chosen to optimise resolution of the
vertically evolving disc, but minor variations are largely
inconsequential as the systems evolves from this state. In our models
we have, however, placed at the inner disc edge at the co-rotation
radius, as opposed to the dust destruction radius used in
\cite{walker_2004}. The initial disc scaleheight at 100\,AU, $h(100)$,
was set to 25\,AU. As the simulation used vertical hydrostatic
equilibrium and dust sublimation, both the disc scaleheight and inner
edge location then evolved in the systems dependent on the input
parameters. In this section we discuss the structure of the discs in
terms of these two generated characteristics, i.e the disc scaleheight
and inner edge location.

@@

\begin{figure*}
\begin{center}
  \subfigure[]{\includegraphics[scale=0.4,angle=0]{./Fig/flare_-12_183.ps}\label{flare_-12_183}}
  \subfigure[]{\includegraphics[scale=0.4,angle=0]{./Fig/flare_-7_193.ps}\label{flare_-7_193}}
\end{center}
 \caption{XXXNJM REMOVE WHITESPACEXXXFigure showing the density structure ($\log \rho$) of the
   BDD system with $M_*=0.04M_{\odot}$, Age=1 Myrs, $\tau$=5, areal
   coverage=10\% and accretion rate of (a) \logmdot=$-12$ and (b) \logmdot = $-7$.}
\end{figure*}

\cite{walker_2004} state that the degree of disc flaring depends on
the disc temperature structure and the mass of the central star, with
the disc scaleheight $h\propto (\frac{T_{\rm disc}}{M_*})^{1/2}$
\citep{shakura_1973}. Recently however \cite{ercolano_2009} proposed
the inverse relation of flaring with stellar mass, i.e. $h\propto
M_*$. This suggestion was based on evidence from \cite{allers_2006},
where SEDs for 17 systems in the mass range 6$M_{\rm
  Jup}<M_*<$350$M_{\rm Jup}$ were fit with flared or flat disc
models. In general, \cite{allers_2006} find that lower mass objects
achieve better fits with the flat disc models and higher mass objects
with the flared discs.

The results of \cite{allers_2006} show that above a mass of 50$M_{\rm
  Jup}$ all objects (6/17) are better fit with flared discs. Whilst at
masses below 50$M_{\rm Jup}$ only one object is better fit by the
flared disc model, with the remaining objects (10/17) better fit with
flat models. Whether, this result is statistically significant enough
to assert a $h\propto M_*$ is doubtful as the fitting process contains
(presumably) two fixed scaleheight distributions. Therefore, for our
study we continue to assume that our flared BDD systems will have
larger characteristic scaleheights than typical CTTS systems.

A  comparison of Figures \ref{flare_-12_183} and
\ref{flare_-7_193} shows an increase in the scaleheight at 50\,AU of
$>$5 AU, as the accretion rate moves from \logmdot =$-12$ to
$-7$. However, despite this small change with high levels of accretion
our grid shows scaleheights comparable to the work of
\cite{walker_2004} and as such result in similar consequences for the
SEDs and photometric magnitudes. The effects of this flaring and the
increase in flaring for very high accretion levels on SEDs and
photometric magnitudes are discussed in Sections \ref{observables}
and \ref{isochrones}, respectively.

@@

Disentangling the accretion and stellar photospheric flux, in order to
derive accretion rates becomes difficult as $T_{\rm acc} \to T_{\rm
  eff}$, which for an accretion rate of \logmdot = $-9$, occurs at a
fractional coverages of 20 and 10\%, for rotation periods of 5 and 0.5
days (with $M_*=0.04M_{\odot}$ and an age of 1 Myr). Additionally,
higher accretion rates can produce enough flux to `veil' the
underlying photospheric features, and significantly change the peak
flux levels. Heavy veiling of atomic bands has been observed in CTTS
systems \citep{kenyon_1995}.

Figure \ref{acc_dom} shows the affect of increasing the accretion
blackbody flux (for increasing accretion rates) for a BD star,
$M=$0.04$M_{\odot}$, at 1 Myr. Whilst the photons originating from the
star (both from $L_*$ and $L_{\rm acc}$) will be tagged as stellar by
{\sc torus} we can separate these flux contributions simply by
observing the naked star system. The panels in Figure \ref{acc_dom}
show the flux from a naked system, with no treatment of the disc. This
enables us to view the effect of increasing accretion rate on the
photospheric flux in isolation. The accretion rates included in all
panels are \logmdot = $-8$, $-9$ and $-12$ (blue, black and red lines
respectively). The bottom panels show the systems with a
rotation period of 5 days and top panels for those with a
rotation period of 0.5 days. Given our assumption that accretion
occurs from the co-rotation radius, decreasing the rotational period
moves this accretion radius closer to the star, $R_{\rm inner}\propto
\tau^{2/3}$ (see Equation \ref{inner_eq}). As the accretion radius
moves father from the star the potential energy released by the
accreted material is increased. This effect can be seen
when comparing the top and bottom panels, although
the effect is marginal for all but the highest displayed accretion
rates.

@@

XXXTHIS SECTION IS ON THE INNER EDGE AND NEEDS SORTINGXXX

Figure \ref{inner_inc} shows the SEDs for the viewing angles of 0, 27,
39, and 48 $^{\circ}$, close to face on. The top and bottom panels of
Figure \ref{inner_inc} show SEDs for models with a curved, as
exemplified by Figure \ref{flare_-7_193}, and a flat inner wall, as
shown in Figure \ref{flare_-12_183} respectively. The specific models
chosen have been selected to minimise flux differences between the two
SEDs but have distinct inner edges. The top panel has a mass of
0.04$M_{\odot}$, \logmdot = -8, areal coverage of 1\%, a rotational
period of 5 days and an age of 10 Myrs. The bottom panel has a the
same model but with rotation period of 0.5 days and an age of 1
Myrs. Comparing the top and bottom panel show that flux levels for the
higher accretion rate, with the larger and curved inner edge, vary
less distinctly with inclination angle than the systems with a
vertical wall. Although this effect is marginal. The weakness of the
effect is probably due to mismatches in the disc structure and flux
levels. Ideally, we would take the curved inner edge model and replace
the inner edge with a vertical wall.

@@

XXXEVERYTHING IS FINE FROM HEREXXX

@@

To delineate the effects of the accretion rate and circumstellar discs
we have subdivided the grid into two groups (as discussed in Sections
\ref{disc_struct}), those with accretion rates typical for higher mass
CTTS objects, defined as \logmdot = $-$12 (negligible) to \logmdot = $-$9 and
those with elevated accretion rates, where \logmdot $>-$9. For several of
the plots in this section the magnitude and colours for the
$M_*=0.01M_{\odot}$ systems at high inclinations become extremely
faint and red. In some cases these objects appear slightly without the
scale of the diagram the axes were limited in this way to be able to
show the changes in colour and magnitude for the majority of stars
better. The colours and magnitudes for these lowest mass stars are not
necessarily unreliable but simply hinder the aesthetics of the plots,
and as they are at the limits of our grid we have decided to omit them
from some Figures by trimming the axes. Additionally, stars at the
highest inclinations, i.e edge on disc systems, are often omitted from
the Figures due to their extremely faint magnitudes, meaning they
would not be practically observable.

@@

The use of pre-MS isochrones for the derivation of single star
parameters is at the present time not proven to be reliable \citep[see
discussion in][]{mayne_2008}. Practically, therefore, median
ages are derived from populations. Subsequently, derived masses are
still unreliable but at least based on a consistent age. This problem
is being addressed by Bell et al (in prep), where $K$ band photometry
and known eclipsing binaries are being used to refine pre-MS
isochrones. In this section we plot the data for our 1 Myr systems
only and explore the resulting scatters caused by the disc presence
and accretion luminosity.

@@

As can be seen in Figures \ref{age_mass_VVI} and \ref{age_mass_JJK}
current accretion and disc presence in a star and disc systems creates
a scatter in our simulated photometry indicative of a much larger
isochronal age or mass spread. Indeed, for many BDD systems even at
nominal accretion rates of \logmdot = $-$11 or $-$12, for our
simulations, the colours of these stars move significantly blueward of
the expected BD locus in a $M_V$, $(V-I)_0$ CMD and redward in a
$M_J$, $(J-K)_0$ CMD. As, in the case of typical accretors, the input
variables are in the range expected for a BD population, one could
reasonably expect observed true BD locii to show a similar scatter. It
is clear that even a wide photometric selection would not include all
of, even the negligibly accreting systems at expected
inclinations. Whilst the derivation of masses and ages for these
objects will be difficult. Furthermore, for the higher accretion rates
of \logmdot = $-$9 or $-$10 the movement of the star within the CMD
will effectively move the star into the contamination region expected
for background CTTS or MS stars at a $(V-I)_0$ of $\leq$1.5 and $M_V$
12-10, and as such the star would not be included in a photometrically
selected BD sample. The solid green line, in the top panels of Figures
\ref{age_mass_VVI} and \ref{age_mass_JJK}, showing the 1 Myr isochrone
of \cite{siess_2000} at a distance of 250 pc and extinction of $A_V=2$
mags shows that the BDD systems with higher accretion rates could
easily be confused for a background CTTS or MS population. Indeed miss
classification of a BDD system as a CTTS system has already been
revealed in \cite{white_2003}. For the extreme accretors, as shown in
the right panels of Figures \ref{age_mass_VVI} and \ref{age_mass_JJK},
there is little chance of these objects being classed,
photometrically, as BD candidates or being assigned the correct
mass. Therefore, if one attempted to locate a population of BDD
systems with elevated or extreme accretion rates, target selection
would have to be placed at much brighter magnitudes, and bluer for
optical or redder for IR, colours.

This scatter for both typical and extreme accretors is a strong
function of inclination, where, as the inclination is increased the
objects are pushed lower in the CMD. Indeed, for the edge on cases
some objects have magnitudes fainter than those shown (for instance
$M_V\approx 20$). This is expected as the star becomes obscured by the
flared disc, interestingly for typical systems the bottom right panels
show that this occurs for inclinations above around 71$^{\circ}$ (as
found in Figure \ref{flare_inc}) in most cases.  However, even for
the lower inclination angles some objects have very faint magnitudes,
this is due to the disc flaring leading to a smaller opening angle and
is discussed in Section \ref{disc_struct}. Crucially, the top panels
show that the scatter from the isochrone is, generally, correlated
with accretion rate. Effectively, as the accretion rate increases the
BDD system moves farther away from the isochrone and is therefore less
likely to be classified as a BDD system and included in any target
samples of such objects. Overall, the dominant scattering effect for
BDD systems in an optical CMD appears to be caused by accretion rate
and inclination, and therefore obscuration effects of the disc on the
star. This suggests that for a given photometric survey of BDD systems
accreting at typical accretion rates and with an expected range of
inclinations (centred on around 60$^{\circ}$) one would expect to
exclude a significant fraction of these objects from any isochrone
based selection.

@@

Therefore, if one adopts the range of input parameters we have used
(see Section \ref{par_space} for justification), our simulated
photometry shows, qualitatively, that a coeval 1 Myr population of
accreting BD stars and BDD systems, with typical accretion rates and
range of inclinations, will exhibit a significant scatter in apparent
isochronal age. Furthermore, objects with typical (and extreme)
accretion rates are scattered sufficiently in CMD space to prohibit
their identification as pre-MS BDs. Indeed, these objects would not be
included in a photometrically selected sample of BDs, and as such are
unlikely to be assigned the correct masses or ages. The scatter from
the naked 1 Myr systems, generally, increases with increasing
accretion rate. It is important to note at this point that these
conclusions are qualitative, and obviously based on our
assumptions. However, the repercussion for isochronal age derivation
and sample selection in the BD regime could be profound. Indeed this
study has only included the effects of current or ongoing accretion
from a disc, it has neglected any effects of accretion, both past and
present, on the evolution of the central star. This past accretion
could also act to reduce the stars radius, accelerating contraction
\citep{tout_1999,siess_1999} and introducing additional scatter in a
coeval population proportional to the range in accretion rates
\citep[see][for full discussion]{mayne_2008}.  Our findings support
and extend those of \cite{walker_2004}, showing that a significant
scatter is caused by obscuration of the central source by the central
disc, which (in our case) is in turn a function of accretion rate. In
addition we have shown that the typical accretion rates may scatter
BDD systems into the region of a CMD occupied by the CTTS or
background MS locus.

@@

We have also modeled the same parameter space using various analytical
disc structures as opposed to vertical hydrostatic
equilibrium. Additionally, we are extending the parameter range of our
grid. Finally, we have created an online fitting tool. All of which
will be featured in upcoming publications.

\subsection{Summary}
\label{conc_summary}

In summary, our simulated dataset shows that for typical parameter
ranges for BD stars and BDD systems, disc presence and accretion flux
lead to:

Difficulty deriving the following stellar parameters for a coeval
population:

>>>>

\begin{abstract}

  We present a grid of simulated accreting brown dwarf systems with
  (and without) circumstellar discs. These simulations include solving
  for vertical hydrostatic and radiative equilibirum
  self-consistently, and a sophisticated implementation of dust
  sublimation. We have simulated observations of the spectral energy
  distributions and using several broadband photometric systems. The
  resulting magnitudes and spectra are publicly available.

  Analysis of the disc structures and simulated observations reveal a
  natural dichotomy in accretion rates, with \logmdot $>-$9 and $\leq
  -$9 classed as extreme and typical accretors respectively. The
  photospheric emissions of the extreme accretors are shown to be
  heavily veiled, and significant sublimation of the dust disc is
  apparent in these systems.

  Derivation of ages and masses from our simulated photometry, using
  isochrones, is demostrated to be unreliable even for typical
  accretors. Current brown dwarf disc candidate selection criteria
  have been shown to be largely reliable when applied to our model
  grid, for both accretion regimes. However, we suggest improved
  selection criteria in several colour indices, derived from our grid.

  We also show that as accretion rates increase, due to photometric
  scatter, brown dwarf disc systems are less likely to be correctly
  identified. This suggests that, within our grid, systems with higher
  accretion rates would be preferentially lost during brown dwarf
  target selection. Therefore, we suggest that observations used to
  assert a $\dot{M}\propto M_*^{\sim 2}$ relationship may contain an
  intrinsic selection bias.

\end{abstract}

@@

Comparison of accretion rates across the pre-main-sequence (PMS) mass
spectrum has indicated that the accretion rate is strongly correlated
with PMS mass, \citep{muzerolle_2003,natta_2004,natta_2006}, with an
approximate form $\dot{M}_{\rm acc}\propto M^{\sim 2}_{*}$. Since in
the canonical picture the accretion rate is driven by the disc
viscosity, and should be indepedent of the mass of the central object,
this correlation is somewhat surprising. There is some danger that the
correlation is the result of, or at least strengthened by, the
presence of observational biases \citep{clarke_2006}: At the high-mass
(CTTS) end of the mass spectrum the lowest accretion-rates cannot be
measured via continuum methods since the excess is too small in
contrast with the photopsheric emission. The emission lines will also
be weak, and indeed the H$\alpha$ equivalent width (EW) may be less
than the 10\AA\ that traditionally demarcates classical from
weak-lined T Tauri stars. Such objects should show doppler-broadened
profiles, but the line wings will be weak, and the presence of
underlying H$\alpha$ absorption may become dominant. Thus there may be
a population of low-accretion-rate CTTS which are current missing from
the surveys.

@@

Nonetheless it is worth considering the observed effect that mass
accretion at typical CTTS rates onto a BD might have. Parity between
the BD photospheric luminosity and the accretion luminosity occurs at
relatively low accretion rates \citep{clarke_2006} and such objects
may not be detected as BDs at all in photometric surveys. Even before
such an extreme case is reached, the additional luminosity provided by
the accretion should have measurable effect on the BD
colours. Furthermore the additional flux will be reprocessed by the
disc, altering its scaleheight and possibly the shape of its inner
edge.

@@

Little observational evidence can be found for approximate sizes of
accretion hot spots due to their more transient nature and often
smaller coverages, when compared to cooler or `plage' spots
\citep{herbst_2007}. \cite{bouvier_1995} modeled the
size of the cool spots on solar-type stars for a selection of
periodically variable candidates. They found projections of cooler
spots, onto the stellar disc, of a few to $\sim$60\%.
\cite{bouvier_1995} also found projected sizes, onto the
stellar disc, of typically a few \% to around 10\% for hot spots.
\cite{bertout_1996} used observations of YY Orionis monitoring
flux amplitude variations as a function of wavelength to derive a
probable hot spot area of around 10\%. The spot temperature was also
modeled for YY Orionis in \cite{bertout_1996}, resulting in a
best fitting areal coverage of 11\%. Therefore, to bound the probable
areal coverage range of the accretion hot spots we have adopted areal
coverages of 1 and 10\%.

@@

Accretion rates derived for pre-MS stars are of order \logmdot = $-6$
to $-11$ \citep{natta_2006}, with the largest accretion rates found in
so-called FU Orionis type objects. For the more typical accretion
rates \citep[$\dot{M}=$10$^{-11}$ to 10$^{-8.9} M_{\odot}yr^{-1}$, for
  TTS,][]{dahm_2009}, several studies have now suggested that the
accretion rate is strongly correlated with the mass of the central
star. This relationship was perhaps first suggested by
\cite{muzerolle_2003} using various accretion diagnostics. Later,
\citep{muzerolle_2005} derived a relationship of approximately
$\dot{M}\propto M_*^{\sim 2}$. Further evidence was put forward by
\cite{natta_2004}, where accretion rates as low as 5$\times$
10$^{-12}M_{\odot} yr^{-1}$ were found for BD stars. More recently,
even lower accretion rates of $\approx$10$^{-13}M_{\odot} yr^{-1}$,
have been derived for BDs by \cite{herczeg_2009}.  Further support for
a dependence of accretion rate on stellar mass was apparent in the
significantly more homogeneous dataset of
\cite{natta_2006}. \cite{natta_2006} analysed a set of accretion rates
and masses derived for pre-MS stars in $\rho$ Ophiuchi and compared
these results to stars in Taurus. They found that the accretion rate
scales with central object mass into the BD regime, although with
significant scatter.

As the relationship $\dot{M}\propto M_*^{\sim 2}$ predicts lower accretion
rates for BD mass objects it is essential that we model systems at
higher accretion rates, which may have been missed in current
observational studies. Therefore, we have adopted accretion rates of
\logmdot = $-6$, $-7$, $-8$, $-9$, $-10$, $-11$ \& $-12$.

@@

The disc surface density is conserved in the initial density
structure, $\Sigma ^{\beta - \alpha}$. The initial scaleheight ($h$)
and density ($\rho$) follow $h\propto r^{\beta}$
($h=h_0(r/R_*)^{\beta}$) and $\rho \propto r^{-\alpha}$ ($\rho
=\rho_0\frac{R_*}{r}exp(-\frac{1}{2}[z/h(r)]^2)$) respectively., where
$r$ and $z$ are the radial and vertical coordinates. The initial
values of $\alpha$ and $\beta$ were 2.1 and 1.1 respectively. The
values for $\alpha$ and $\beta$ were chosen to optimise resolution of
the vertically evolving disc, but minor variations are largely
inconsequential as the systems evolves from this state. In our models
we have, however, placed at the inner disc edge at the co-rotation
radius, as opposed to the dust destruction radius used in
\cite{walker_2004}. The initial disc scaleheight at 100\,AU, $h(100)$,
was set to 25\,AU. As the simulation used vertical hydrostatic
equilibrium and dust sublimation, both the disc scaleheight and inner
edge location then evolved in the systems dependent on the input
parameters. In this section we discuss the structure of the discs in
terms of these two generated characteristics, i.e the disc scaleheight
and inner edge location.

@@

\begin{figure*}
\begin{center}
  \subfigure[]{\includegraphics[scale=0.4,angle=0]{./Fig/flare_-12_183.ps}\label{flare_-12_183}}
  \subfigure[]{\includegraphics[scale=0.4,angle=0]{./Fig/flare_-7_193.ps}\label{flare_-7_193}}
\end{center}
 \caption{Figure showing the density structure ($\log \rho$) of the
   BDD system with $M_*=0.04M_{\odot}$, Age=1 Myrs, $\tau$=5, areal
   coverage=10\% and accretion rate of (a) \logmdot=$-12$ and (b) \logmdot = $-7$.}
\end{figure*}

\cite{walker_2004} state that the degree of disc flaring depends on
the disc temperature structure and the mass of the central star, with
the disc scaleheight $h\propto (T_{\rm disc}/M_*)^{1/2}$
\citep{shakura_1973}. Recently however \cite{ercolano_2009} proposed
the inverse relation of flaring with stellar mass, i.e. $h\propto
M_*$. This suggestion was based on evidence from \cite{allers_2006},
where SEDs for 17 systems in the mass range 6$M_{\rm
  Jup}<M_*<$350$M_{\rm Jup}$ were fit with flared or flat disc
models. In general, \cite{allers_2006} find that lower mass objects
achieve better fits with the flat disc models and higher mass objects
with the flared discs.

The results of \cite{allers_2006} show that above a mass of 50$M_{\rm
  Jup}$ all objects (6/17) are better fit with flared discs. Whilst at
masses below 50$M_{\rm Jup}$ only one object is better fit by the
flared disc model, with the remaining objects (10/17) better fit with
flat models. Whether this result is statistically significant enough
to assert a $h\propto M_*$ is doubtful as the fitting process contains
(presumably) two fixed scaleheight distributions. Therefore, for our
study we continue to assume that our flared BDD systems will have
larger characteristic scaleheights than typical CTTS systems.

A comparison of Figures \ref{flare_-12_183} and \ref{flare_-7_193}
shows an increase in the scaleheight at 50\,AU of $>$5 AU, as the
accretion rate moves from \logmdot =$-12$ to $-7$. However, despite
this small change with high levels of accretion our grid shows
scaleheights comparable to the work of \cite{walker_2004} and as such
result in similar consequences for the SEDs and photometric
magnitudes. The effects of this flaring and the increase in flaring
for very high accretion levels on SEDs and photometric magnitudes are
discussed in Sections \ref{observables} and \ref{isochrones}
respectively.

@@

Disentangling the accretion and stellar photospheric flux, in order to
derive accretion rates becomes difficult as $T_{\rm acc} \to T_{\rm
  eff}$, which for an accretion rate of \logmdot = $-9$, occurs at a
fractional coverages of 20 and 10\%, for rotation periods of 5 and 0.5
days (with $M_*=0.04M_{\odot}$ and an age of 1 Myr). Additionally,
higher accretion rates can produce enough flux to `veil' the
underlying photospheric features, and significantly change the peak
flux levels. Heavy veiling of atomic bands has been extensively
observed in CTTS systems \citep[e.g][]{kenyon_1995}.

Figure \ref{acc_dom} shows the affect of increasing the accretion
blackbody flux (for increasing accretion rates) for a BD star,
$M=$0.04$M_{\odot}$, at 1 Myr. Whilst the photon packets originating
from the star (both from $L_*$ and $L_{\rm acc}$) will be tagged as
stellar by {\sc torus} we can separate these flux contributions simply
by observing the naked star system. The panels in Figure \ref{acc_dom}
show the flux from a naked system, with no treatment of the disc. This
enables us to view the effect of increasing accretion rate on the
photospheric flux in isolation. The accretion rates included in all
panels are \logmdot = $-8$, $-9$ and $-12$ (blue, black and red lines
respectively). The bottom panels show the systems with a rotation
period of 5 days and top panels for those with a rotation period of
0.5 days. Given our assumption that accretion occurs from the
co-rotation radius, decreasing the rotational period moves this
accretion radius closer to the star, $R_{\rm inner}\propto \tau^{2/3}$
(see Equation \ref{inner_eq}). As the accretion radius moves father
from the star the potential energy released by the accreted material
is increased. This effect can be seen when comparing the top and
bottom panels, although the effect is marginal for all but the highest
displayed accretion rates.

@@

Figure \ref{inner_inc} shows the SEDs for the viewing angles of 0, 27,
39, and 48 $^{\circ}$, close to face on. The top and bottom panels of
Figure \ref{inner_inc} show SEDs for models with a curved, as
exemplified by Figure \ref{flare_-7_193}, and a flat inner wall, as
shown in Figure \ref{flare_-12_183} respectively. The specific models
chosen have been selected to minimise flux differences between the two
SEDs but have distinct inner edges. The top panel has a mass of
0.04$M_{\odot}$, \logmdot = -8, areal coverage of 1\%, a rotational
period of 5 days and an age of 10 Myrs. The bottom panel has a the
same model but with rotation period of 0.5 days and an age of 1
Myrs. Comparing the top and bottom panel show that flux levels for the
higher accretion rate, with the larger and curved inner edge, vary
less distinctly with inclination angle than the systems with a
vertical wall. The difference in inclination dependence of the fluxes
between the curved and vertical inner walls is marginal. This is due
to differences in the stellar flux levels and disc
structures. Essentially we are not comparing like for like systems,
ideally we would replace the curved inner edge with a vertical wall in
an eample system and recalculate the temperature structure and SED.

@@

@@

To delineate the effects of the accretion rate and circumstellar discs
we have subdivided the grid into two groups (as discussed in Sections
\ref{disc_struct}), those with accretion rates typical for higher mass
CTTS objects, defined as \logmdot = $-$12 (negligible) to \logmdot =
$-$9 and those with elevated accretion rates, where \logmdot
$>-$9. For several of the plots in this section the magnitude and
colours for the $M_*=0.01M_{\odot}$ systems at high inclinations
become extremely faint and red. As we are interested in the morphology
of the population as a whole these systems are sometimes without the
scale of the CMDs and CoCoDs. Additionally, stars at the highest
inclinations, i.e edge on disc systems, are often omitted from the
Figures due to their extremely faint magnitudes, meaning they would
not be practically observable.

@@

The use of pre-MS isochrones for the derivation of single star
parameters is at present not proven to be reliable \citep[see
discussion in][]{mayne_2008}. Practically, therefore, median ages are
derived from populations. Subsequently, derived masses are still
unreliable but at least based on a consistent age. This problem is
being addressed by Bell et al (in prep), where $K$ band photometry and
known eclipsing binaries are being used to refine pre-MS
isochrones. In this section we plot the data for our 1 Myr systems
only and explore the resulting scatters caused by the disc presence
and accretion luminosity.

@@

As can be seen in Figures \ref{age_mass_VVI} and \ref{age_mass_JJK}
current accretion and disc presence in a star and disc systems creates
a scatter in our simulated photometry indicative of a much larger
isochronal age or mass spread. Indeed, for many BDD systems even at
nominal accretion rates of \logmdot = $-$11 or $-$12, for our
simulations, the colours of these stars move significantly blueward of
the expected BD locus in a $M_V$, $(V-I)_0$ CMD and redward in a
$M_J$, $(J-K)_0$ CMD. As, in the case of typical accretors, the input
variables are in the range expected for a BD population, one could
reasonably expect observed true BD locii to show a similar scatter. It
is clear that even a wide photometric selection would not even include
all of the negligibly accreting systems at expected
inclinations. Whilst the derivation of masses and ages for these
objects will be difficult. Furthermore, for the higher accretion rates
of \logmdot = $-$9 or $-$10 the movement of the star within the CMD
will effectively move the star into the contamination region expected
for background CTTS or MS stars at a $(V-I)_0$ of $\leq$1.5 and $M_V$
12--10, and as such the star would not be included in a
photometrically selected BD sample. The solid green line, in the top
panels of Figures \ref{age_mass_VVI} and \ref{age_mass_JJK}, showing
the 1 Myr isochrone of \cite{siess_2000} at a distance of 250 pc and
extinction of $A_V=2$ mags shows that the BDD systems with higher
accretion rates could easily be confused for a background CTTS or MS
population. Indeed misclassification of a BDD system as a CTTS system
has already been revealed in \cite{white_2003}. For the extreme
accretors, as shown in the right panels of Figures \ref{age_mass_VVI}
and \ref{age_mass_JJK}, there is little chance of these objects being
classed, photometrically, as BD candidates or being assigned the
correct mass. Therefore, if one attempted to locate a population of
BDD systems with elevated or extreme accretion rates, target selection
would have to be placed at much brighter magnitudes, and bluer for
optical or redder for IR, colours.

This scatter for both typical and extreme accretors is a strong
function of inclination, where, as the inclination is increased the
objects are pushed lower in the CMD. Indeed, for the edge on cases
some objects have magnitudes fainter than those shown (for instance
$M_V\approx 20$). This is expected as the star becomes obscured by the
flared disc, interestingly for typical systems the bottom right panels
show that this occurs for inclinations above around 71$^{\circ}$ (as
found in Figure \ref{flare_inc}) in most cases.  However, even for the
lower inclination angles some objects have very faint magnitudes, this
is due to the disc flaring leading to earlier obscuration of the star
as discussed in Section \ref{disc_struct}. Crucially, the top panels
show that the scatter from the isochrone is, generally, correlated
with accretion rate. Effectively, as the accretion rate increases the
BDD system moves farther away from the isochrone and is therefore less
likely to be classified as a BDD system and included in any target
samples of such objects. Overall, the dominant scattering effect for
BDD systems in an optical CMD appears to be caused by accretion rate
and inclination, and therefore obscuration effects of the disc on the
star.

@@

Therefore, if one adopts the range of input parameters we have used
(see Section \ref{par_space} for justification), our simulated
photometry shows, qualitatively, that a coeval 1 Myr population of
accreting BD stars and BDD systems, with typical accretion rates and
range of inclinations, will exhibit a significant scatter in apparent
isochronal age. Furthermore, objects with typical (and extreme)
accretion rates are scattered sufficiently in CMD space to prohibit
their identification as pre-MS BDs. Indeed, these objects would not be
included in a photometrically selected sample of BDs, and as such are
unlikely to be assigned the correct masses or ages. The scatter from
the naked 1 Myr systems, generally, increases with increasing
accretion rate. It is important to note at this point that these
conclusions are qualitative, and obviously based on our
assumptions. However, the repercussion for isochronal age derivation
and sample selection in the BD regime could be profound. Indeed this
study has only included the effects of current or ongoing accretion
from a disc, it has neglected any effects of accretion, both past and
present, on the evolution of the central star. This past accretion
could also act to reduce the stars radius, accelerating contraction
\citep{tout_1999,siess_1999} and introducing additional scatter in a
coeval population proportional to the range in accretion rates
\citep[see][for full discussion]{mayne_2008}. In addition we have
shown that the typical accretion rates may scatter BDD systems into
the region of a CMD occupied by the CTTS or background MS locus.

@@

We have also modeled the same parameter space using various analytical
disc structures as opposed to vertical hydrostatic
equilibrium. Additionally, we are extending the parameter range of our
grid. Finally, we have created an online fitting tool. All of which
will be featured in upcoming publications. Finally, we intend to
explore the efficacy of accretion rate diagnostics by simlating line
profiles of accreting systems for a range of accretion rates.
